---
title: "p8105_hw6_yw3773"
output: github_document
---
```{r}
library(tidyverse)
library(modelr)
library(mgcv)
```

## Problem 1

#### data cleaning

```{r}
## data cleaning, convert numeric to factor, there is no missing value
birth_df = read_csv("birthweight.csv")

birth_df = birth_df %>%
  mutate(
    babysex = as.factor(babysex),
    babysex = recode(babysex, `1` = "male", `2` = "female"),
    frace = as.factor(frace),
    frace = recode(frace, `1` = "White", `2` = "Black", `3` = "Asian", `4` = "Puerto Rican", 
                   `8` = "Other", `9` = "Unknown"),
    malform = as.factor(malform),
    malform = recode(malform, `0` = "absent", `1` = "present"),
    mrace = as.factor(mrace),
    mrace = recode(mrace, `1` = "White", `2` = "Black", `3` = "Asian", `4` = "Puerto Rican", 
                   `8` = "Other")
  )
```

This data set contains `r ncol(birth_df)` variables and `r nrow(birth_df)` observations. I converted "babysex", "frace" (father's race), "malform" (presence of malformations), "mrace" (mother's race) into factor variables with appropriate names according to the variables' information. There is no missing value.

#### fitting my first model

```{r}
## hypothesis plot for model driven
birth_df %>%
  ggplot(aes(y = bwt, x = bhead)) + 
  geom_point(alpha = 0.5, size = 0.5) + 
  labs(title = "Baby's birth weight vs. the head circumference at birth")

birth_df %>% 
  ggplot(aes(y = bwt, x = babysex)) + geom_boxplot() +
  labs(title = "Baby's birth weight with respect to their birth sex")

```

I was quite interested in how baby's head circumference at birth and the birth sex are associated with their birth weight. Then I plot two graphs as above, from the first graph, it seems like there is a positive relationship between the head circumference and birth weight. From the second graph, although the two box plots are overlapped, the male's median is a little bit higher than the female's median. So I decide to fit a linear regression of birth weight on head circumference at birth and birth sex, and I assume there is no interactions. 

```{r}
## fit my model for birth weight
fit_1 = lm(bwt ~ babysex + bhead, data = birth_df)
fit_1 %>% 
  broom::tidy() 

```

```{r}
birth_df %>%
  modelr::add_predictions(fit_1) %>%
  modelr::add_residuals(fit_1) %>%
  ggplot(aes(y = resid, x = pred)) + 
  geom_point(alpha = 0.3, size = 1) +
  labs(title = "Model residuals against fitted values") +
  xlab("fitted value") + ylab("model residual")
```

From the model results we can see that, 1cm increase in baby's head circumference is associated with 238.3g increase in baby's weight. Also, being a female has a positive effect on baby's birth weight comparing to male. From the plot of model residuals against fitted values, we can see that all points are randomly distributed abound 0, except a few outliers.

#### fitting the second and third model, and comparing them with model 1

```{r}
fit_2 = lm(bwt ~ blength + gaweeks, data = birth_df)
fit_2 %>% broom::tidy()

fit_3 = lm(bwt ~ bhead + blength + babysex + bhead*blength*babysex, data = birth_df)
fit_3 %>% broom::tidy()
```

```{r}
birth_df =
  crossv_mc(birth_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

birth_df = birth_df %>%
  mutate(
    mod_1 = map(train, ~fit_1, data = .x),
    mod_2 = map(train, ~fit_2, data = .x),
    mod_3 = map(train, ~fit_3, data = .x)
  ) %>%
  mutate(
    rmse_mod_1 = map2_dbl(mod_1, test, ~rmse(model = .x, data = .y)),
    rmse_mod_2 = map2_dbl(mod_2, test, ~rmse(model = .x, data = .y)),
    rmse_mod_3 = map2_dbl(mod_3, test, ~rmse(model = .x, data = .y))
  )

```

```{r}
birth_df %>%
  select(starts_with("rmse")) %>%
  pivot_longer(
    everything(),
    names_to = "model", values_to = "rmse",
    names_prefix = "rmse_"
  ) %>%
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  labs(title = "Prediction errors for the three models")
```

```{r}
birth_df %>%
  select(starts_with("rmse")) %>%
  pivot_longer(
    everything(),
    names_to = "model", values_to = "rmse",
    names_prefix = "rmse_"
  ) %>%
  group_by(model) %>%
  summarize(avg_rmse = mean(rmse))
```

After fitting the second and third model and comparing them with the first model I fitted. From both the violin plot and the mean prediction errors, we can see that the third model (using head circumference, length, sex, and all interactions) is better fitted comparing two the first and second models, as it had lower rmse.

## Problem 2

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r}
weather_results = weather_df %>%
  bootstrap(100, id = "strap_number") %>%
  mutate(
    models = map(.x = strap, ~lm(tmax ~ tmin, data = .x)),
    results = map(models, broom::tidy),
    results_2 = map(models, broom::glance)
  ) %>%
  select(strap_number, results, results_2) %>%
  unnest(results,results_2)

weather_results
```

```{r}
plot_df = weather_results %>%
  select(strap_number, term, estimate) %>%
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) %>%
  mutate(intercept = `(Intercept)`) %>%
  select(-`(Intercept)`) %>%
  mutate(log_beta = log(intercept * tmin))

plot_df %>%
  ggplot(aes(x = log_beta)) + geom_density()

plot_df %>%
  summarize(
    ci_lower_log_beta = quantile(log_beta, 0.025), 
    ci_upper_log_beta = quantile(log_beta, 0.975))

```

```{r}
weather_results %>%
  filter(term == "tmin") %>%
  ggplot(aes(x = adj.r.squared)) + geom_density()

weather_results %>%
  filter(term == "tmin") %>%
  summarize(
    ci_lower_r = quantile(adj.r.squared, 0.025),
    ci_upper_r = quantile(adj.r.squared, 0.975)
  )

```













